---
title: "Working with rasters and LANDIS-II"
author: "Sam Flake"
date: "8/16/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## LANDIS-II

LANDIS-II is a flexible landscape simulation model, widely used for improving understanding and forecasting forested landscapes. 

LANDIS-II takes two kinds of inputs: .txt files which give the parameters for the model, and raster files which describe the landscape -- the initial communities, the abiotic conditions, fuels, topography, and other important variables. While rasters are an intuitive way to store spatial data, working with rasters isn't always straightforward. 

This vignette doesn't require LANDIS-II to be installed, but it will present rasters used in LANDIS-II. See xx for details on the model.

Rasters for LANDIS-II have a few general requirements and best practices:

#TODO list these

#TODO create some example rasters, use NECN examples

In this vignette, we will tackle a seemingly simple problem: taking a LANDIS-II landscape and creating a smaller subset of the landscape. Sometimes this is needed to reduce the processing time for a LANDIS-II run, to allow faster iterations when calibrating the model. But this process brings up some important potential pitfalls we need to avoid. 

```{r cars}
summary(cars)
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
